fGFPCA dynamic prediction

### 2023-02-01 f-GFPCA.Rmd
hand code laplace approximation
Laplace and MLE are actually the same thing

@@

```{r, results='hide'}
source(here("Code/OutSampMLE.R"))
source(here("Code/OutSampBayes.R"))
```

@@

```{r, results='hide'}
out_pred_lap_subj1 <- out_pred_clean_laplace(1, mid) %>% 
  left_join(df_true_eta %>% filter(id==1), by = "mid")
out_pred_lap_subj2 <- out_pred_clean_laplace(2, mid) %>% 
  left_join(df_true_eta %>% filter(id==2), by = "mid")
out_pred_lap_subj3 <- out_pred_clean_laplace(3, mid) %>% 
  left_join(df_true_eta %>% filter(id==3), by = "mid")
out_pred_lap_subj4 <- out_pred_clean_laplace(4, mid) %>% 
  left_join(df_true_eta %>% filter(id==4), by = "mid")
```

@@

This method essential finds $\boldsymbol{\xi}$ that maximizes $p(\boldsymbol{Y}|\boldsymbol{\xi})$.

- <span style="color: red;"> Is it equivalent to posterior mean $E(\boldsymbol{\xi}|\boldsymbol{Y})$ under certain conditionss? (according to decision theory) </span>
- <span style="color: red;"> How to use Laplace Approximation to approximate integrals? </span>

https://search.r-project.org/CRAN/refmans/LaplacesDemon/html/LaplaceApproximation.html

@@

```{r, results='hide'}
# true scores:xi

@@

- The three components should be mutally independent? 
- Through some derivation, I think $Var(\boldsymbol{f_0})=\frac{Var(\boldsymbol{Y})}{N} = \frac{1}{N}\boldsymbol{\Phi\Gamma \Phi}+\sigma_{\epsilon}^2\boldsymbol{I}_m$.
- Variance of $\hat{\boldsymbol{\xi}}$ will depend on the method used for its estimation
- $Var(\epsilon) =\hat{\sigma}_{\epsilon}^2$ from FPCA
- Very different approach from in-sample estimation

>>>>

Posterior mean from Laplace method is the same as posterior mode. In this sense, empirical bayes method and MLE are in fact equivalent.

https://search.r-project.org/CRAN/refmans/LaplacesDemon/html/LaplaceApproximation.html
https://bookdown.org/rdpeng/advstatcomp/laplace-approximation.html

@@

```{r, results='hide', message=FALSE}
source(here("Code/OutSampMLE.R"))
source(here("Code/OutSampBayes.R"))
```

@@

```{r, results='hide', message=FALSE}
out_pred_lap_subj1 <- out_pred_clean_laplace(1, mid) %>% 
  left_join(df_true_eta %>% filter(id==1), by = "mid")
out_pred_lap_subj2 <- out_pred_clean_laplace(2, mid) %>% 
  left_join(df_true_eta %>% filter(id==2), by = "mid")
out_pred_lap_subj3 <- out_pred_clean_laplace(3, mid) %>% 
  left_join(df_true_eta %>% filter(id==3), by = "mid")
out_pred_lap_subj4 <- out_pred_clean_laplace(4, mid) %>% 
  left_join(df_true_eta %>% filter(id==4), by = "mid")
```

@@

@@

```{r, results='hide', message=FALSE}
# true scores:xi

@@

- The three components should be mutually independent? 
- Through some derivation, I think $Var(\boldsymbol{f_0})=\frac{Var(\boldsymbol{Y})}{N} = \frac{1}{N}[\boldsymbol{\Phi\Gamma \Phi^T}+\sigma_{\epsilon}^2\boldsymbol{I}_m]$.
- Variance of $\hat{\boldsymbol{\xi}}$ will depend on the method used for its estimation. But isn't the estimated done conditioning on $Var(\boldsymbol{\xi}) = \boldsymbol{\Gamma}$?ß
- $Var(\epsilon) =\hat{\sigma}_{\epsilon}^2$ from FPCA
- Very different approach from in-sample estimation
